\large{\bf Looking forwards and backwards through locally regulated populations }

### 2023-04-17 forwards-and-backwards.tex
.

In this section we verify one of the conditions of the Markov Mapping Theorem,
namely, that ``integrating out levels'' in the generator of the lookdown process
we obtain the generator of the projected process.
We will work with test functions of the form
$$  f(\lp) = \prod_{(x, u) \in \lp} g(x, u) = \exp\left(\langle \log g, \lp \rangle \right) . $$
Furthermore, recall that $\kappa^N(\lp)(\cdot) = \lp(\cdot \times [0, N)) / N$
is the ``spatial projection operator'',
and define the probability kernel $\Gamma^N : \measures \to \mathcal{M}(\IR^d \times [0,N))$
so that for $\eta \in \measures$, if $\hat g(x) = \int_0^N g(x, u) du / N$, then
\begin{align*}
    F^N_g(\eta)
    &:=
    \int f(\lp) \Gamma^N(\eta, d\lp) \\
    &=
    \exp\left(
        N \left\langle
            \log \frac{1}{N} \int_0^N g(x, u) du, \eta(dx)
        \right\rangle
    \right) \\
    &=
    \exp\left(
        N \left\langle \log \hat g(x), \eta(dx) \right\rangle
    \right) ,
\end{align*}
i.e., $\Gamma^N$ assigns independent labels on $[0, N]$ to each of the points in $\eta$.
It follows from Lemma~\ref{def: MP definition of limit} that for test functions of this form
the generator of $\eta^N_t$ is
\begin{align} \label{eqn:pgen_defn_again}
    \begin{split}
    \Pgen^N F^N_g(\eta)
    =
    F^N_g(\eta)
        N \theta & \bigg\langle
        \gamma(x, \eta) \int r(z, \eta) \left( \hat g(z) - 1 \right) q_\theta(x, dz)
    \\ &\qquad \qquad {}
        +
        \mu_\theta(x, \eta) \left( \frac{1}{\hat g(x)} - 1 \right)
        ,
        \eta(dx)
    \bigg\rangle .
    \end{split}
\end{align}
(Note that $f$ here differs from the $f$ used in Lemma~\ref{def: MP definition of limit}
so as to agree with standard usage in the literature on lookdown processes.)
The generator of $\lp^N_t$ is $A^N$, defined in equation~\eqref{eqn:lookdown_generator}.

\comment{TODO: conditions on $f$}

\begin{lemma}
    \label{lem:averaged_generators}
    For all finite counting measures $\eta$ on $\IR^d$,
    \begin{align}
        \int A^N f(\lp) \Gamma^N(\eta, d\lp)
        &=
        \Pgen^N F^N_g(\eta) .
    \end{align}
\end{lemma}

\begin{lemma}
    \label{lem:averaged_generators_limit}
    For all $\eta \in \measures$,
    \begin{align}
        \int A f(\lp) \Gamma(\eta, d\lp)
        &=
        \Pgen^\infty F_g(\eta) .
    \end{align}
\end{lemma}

>>>>

In this section we verify one of the conditions of the Markov Mapping Theorem,
namely, that ``integrating out levels'' in the generator of the lookdown process
we obtain the generator of the projected process.
We will work with test functions of the form
\begin{equation} \label{eqn:f_defn}
    f(\lp) = \prod_{(x, u) \in \lp} g(x, u) = \exp\left(\langle \log g, \lp \rangle \right) ,
\end{equation}
where $0 \le g \le 1$ and $g(x,u) = 1$ for all $u \ge u_g$ for some $u_g < \infty$.
Furthermore, recall that $\kappa^N(\lp)(\cdot) = \lp(\cdot \times [0, N)) / N$
is the ``spatial projection operator'',
and define the probability kernel $\Gamma^N : \measures \to \mathcal{M}(\IR^d \times [0,N))$
so that for $\eta \in \measures$, if $\hat g(x) = \int_0^N g(x, u) du / N$, then
\begin{align*}
    F^N_g(\eta)
    &:=
    \int f(\lp) \Gamma^N(\eta, d\lp) \\
    &=
    \exp\left(
        N \left\langle
            \log \frac{1}{N} \int_0^N g(x, u) du, \eta(dx)
        \right\rangle
    \right) \\
    &=
    \exp\left(
        N \left\langle \log \hat g(x), \eta(dx) \right\rangle
    \right) ,
\end{align*}
i.e., $\Gamma^N$ assigns independent labels on $[0, N]$ to each of the points in $\eta$.
It follows from Lemma~\ref{def: MP definition of limit} that for test functions of this form
the generator of $\eta^N_t$ is
\begin{align} \label{eqn:pgen_defn_again}
    \begin{split}
    \Pgen^N F^N_g(\eta)
    =
    F^N_g(\eta)
        N \theta & \bigg\langle
        \gamma(x, \eta) \int r(z, \eta) \left( \hat g(z) - 1 \right) q_\theta(x, dz)
    \\ &\qquad \qquad {}
        +
        \mu_\theta(x, \eta) \left( \frac{1}{\hat g(x)} - 1 \right)
        ,
        \eta(dx)
    \bigg\rangle .
    \end{split}
\end{align}
(Note that $f$ here differs from the $f$ used in Lemma~\ref{def: MP definition of limit}
so as to agree with standard usage in the literature on lookdown processes.)
The generator of $\lp^N_t$ is $A^N$, defined in equation~\eqref{eqn:lookdown_generator}.

\begin{lemma}
    \label{lem:averaged_generators}
    For all finite counting measures $\eta$ on $\IR^d$,
    if $f$ is of the form~\eqref{eqn:f_defn}, then
    \begin{align}
        \int A^N f(\lp) \Gamma^N(\eta, d\lp)
        &=
        \Pgen^N F^N_g(\eta) .
    \end{align}
\end{lemma}

\begin{lemma}
    \label{lem:averaged_generators_limit}
    For all $\eta \in \measures$,
    if $f$ is of the form~\eqref{eqn:f_defn}, then
    \begin{align}
        \int A f(\lp) \Gamma(\eta, d\lp)
        &=
        \Pgen^\infty F_g(\eta) .
    \end{align}
\end{lemma}
